---
title: "Link Genius Demonstration"
author: "Thomas Henry"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

This is a Demonstration R markdown document for the Link Genius open-source user-friendly record linkage/de-duplication tool. You can run it in VSCode or using "Run Document" in RStudio.

The following chunk contains parameters you may wish to adjust so that it will run the way you want on your computer.

```{r}
port = 8080
```

The following chunk loads up libraries that Link Genius depends on.

```{r}
if( !require(fastLink) ) install.packages("fastLink")
library(fastLink)
 
if( !require(shiny) ) install.packages("shiny")
library(shiny)

#if( !require(reticulate) ) install.packages("reticulate")
#library(reticulate)
#iframe <- import("proxy_frame.iframe", as = "iframe")
#iframe_script <- iframe$create_iframe(port, 500)

#print(iframe_script)
```

The following chunk loads in line the Link Genius logic.  In the future when this project has been funded this will be refactored into an R library.  Until then, this should suffice to demonstrate the functionality. 

```{r}
# Define UI for app that draws a histogram ----
ui <- fluidPage(

  # App title ----
  titlePanel("Hello Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)

# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })
}
```

The following chunk will run the GUI service. If this does not display automatically, please go to the link displayed in your terminal. When you are finished, you can terminate the server by pressing Ctrl-C in the terminal.

```{r}
shinyApp(ui = ui, server = server, options = list(port=port))
```