---
title: "Link Genius Demonstration"
author: "Thomas Henry"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

This is a Demonstration R markdown document for the Link Genius open-source user-friendly record linkage/de-duplication tool. You can run it in VSCode or using "Run Document" in RStudio.

The following chunk contains parameters you may wish to adjust so that it will run the way you want on your computer.

```{r}
port = 8080
```

The following chunk loads up libraries that Link Genius depends on.

```{r}
if( !require(fastLink) ) install.packages("fastLink")
library(fastLink)
 
if( !require(shiny) ) install.packages("shiny")
library(shiny)

if( !require(shinyWidgets) ) install.packages("shinyWidgets")
library(shinyWidgets)

#if( !require(reticulate) ) install.packages("reticulate")
#library(reticulate)
#iframe <- import("proxy_frame.iframe", as = "iframe")
#iframe_script <- iframe$create_iframe(port, 500)

#print(iframe_script)
```

The following chunk creates some defaults.  You do not need to modify these, since the GUI will present these fields.

```{r}
data(samplematch)
# provides:
# dfA
# dfB

varnames <- c("firstname", "middlename", "lastname", "housenum", "streetname", "city", "birthyear")
stringdistMatch <- c("firstname", "middlename", "lastname", "streetname", "city")
partialMatch <- c("firstname", "lastname", "streetname")
```

The following chunk loads in line the Link Genius logic.  In the future when this project has been funded this will be refactored into an R library.  Until then, this should suffice to demonstrate the functionality. 

```{r}
ui <- fluidPage(

  titlePanel("Link Genius Interactive Record Linker"),

  h1("Data Set Summaries"),

  fluidRow(
    column(6, 
      # Input: Select a dataset ----
      h4("Data Set A"),

      # Input: Specify the number of observations to view ----
      numericInput("obsNA", "Number of observations to view:", 10),

      # Include clarifying text ----
      helpText("Note: while the data view will show only the specified",
               "number of observations, the summary will still be based",
               "on the full dataset."),

      h4("Summary"),
      verbatimTextOutput("summaryA"),

      h4("Randomly Sampled Rows"),
      div(tableOutput("viewA"), style = "font-size:50%")
    ),
    column(6,
      # Input: Select a dataset ----
      h4("Data Set B"),

      # Input: Specify the number of observations to view ----
      numericInput("obsNB", "Number of observations to view:", 10),

      # Include clarifying text ----
      helpText("Note: while the data view will show only the specified",
               "number of observations, the summary will still be based",
               "on the full dataset."),

      h4("Summary"),
      verbatimTextOutput("summaryB"),

      h4("Randomly Sampled Rows"),
      div(tableOutput("viewB"), style = "font-size:50%")
    ),

    helpText("Each time this Update View is pressed, another random ",
             "subset of each data set will be displayed."),
    actionButton("updateSummary", "Update View"),

  ),

  h1("Record Link Plan"),

  fluidRow(
    column(2,
      h4("Columns in A"),
      htmlOutput("colsA")
    ),
    column(8,
      div(h3("Columns in A and B"), style = "text-align:center"),
      fluidRow(
          column(4,
            h4("Columns not considered"),
          ),
          column(4,
            h4("Columns to be matched"),
          ),
          column(4,
            h4("Columns to be matched, using partial agreement"),
          )
        ),
      fluidRow(
          column(4,
            selectizeGroupUI(
              inputId = "pickUnused",
              choices = c("1", "2", "3")),
          ),
          column(4,
            pickerInput(
              inputId = "pickMatchers",
              choices = c()),
          ),
          column(4,
            pickerInput(
              inputId = "pickPartial",
              choices = c()),
          )
        )
    ),
    column(2,
      h4("Columns in B"),
      htmlOutput("colsB")
    ),
  )  
)

# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Generate a summary of datasetA ----
  output$summaryA <- renderPrint({
    summary(dfA)
  })

  aNames <- colnames(dfA)
  bNames <- colnames(dfB)

  colsBoth <- Reduce(intersect, list(aNames, bNames))

  output$colsA <- renderText({
      paste(
          "<p>",
          paste(aNames, sep = "</p><p>"),
          "</p>", sep = " "
      )
    })

  output$colsB <- renderText({
      paste(
          "<p>",
          paste(bNames, sep = "</p><p>"),
          "</p>", sep = " "
      )
    })

  # Generate a summary of datasetB ----
  output$summaryB <- renderPrint({
    summary(dfB)
  })

  datasetA <- eventReactive(input$updateSummary, 
      head(sample(dfA), n = isolate(input$obsNA)),
      ignoreNULL = FALSE)

  datasetB <- eventReactive(input$updateSummary, 
      head(sample(dfB), n = isolate(input$obsNB)),
      ignoreNULL = FALSE)

  # Show the first "n" observations of A ----
  # The use of isolate() is necessary because we don't want the table
  # to update whenever input$obs changes (only when the user clicks
  # the action button)
  output$viewA <- renderTable({
    datasetA()
  })

  # Show the first "n" observations of B ----
  # The use of isolate() is necessary because we don't want the table
  # to update whenever input$obs changes (only when the user clicks
  # the action button)
  output$viewB <- renderTable({
    datasetB()
  })

  #observeEvent(input$matchUnused, {
  #    print(input$pickUnused)
      #updatePickerInput(session = session, inputId = "pickUnused",
      #  choices = colsBoth)
  #})
  
}
```

The following chunk will run the GUI service. If this does not display automatically, please go to the link displayed in your terminal. When you are finished, you can terminate the server by pressing Ctrl-C in the terminal.

```{r}
shinyApp(ui = ui, server = server, options = list(port=port))
```